Make data files

### 2021-11-29 make_data.Rmd
Teams and franchises

Processing the files from `data-src` to produce data frames and igraph/network objects for analysis.

# Trades data

Load trades data from Zachary Richardson. Group trades into seasons using April 1 of each year.

```{r}
e <- new.env()
load(here::here("data-src", "data", "NBA_AnalysisData.RData"), envir=e)

trades <- e$trade.df_long |>
  select(-season) |> # This one is wrong
  mutate(
    season_from = case_when(
      month(Date) >= 4 ~ as.numeric(year(Date)),
      month(Date) <= 3 ~ as.numeric(year(Date)) - 1,
      TRUE ~ as.numeric(NA)
    ),
    season_to = season_from + 1,
    season = paste0(season_from, "-", season_to)
  )
```

@@

```{r graphlist}
graphlist <- lapply(
  sort(unique(trades$season)),
  function(s) {
    g <- graph_from_data_frame( 
      edb %>%
        filter(season == s) %>%
        select(P1, P2), 
      directed=FALSE)
    E(g)$weight <- count_multiple(g)
    g <- simplify(g)
    g$season <- s
    g
  }
)
```

>>>>

Processing the files created by Zachary Ricardson (in `data-src`) to produce data frames and igraph objects for analysis.

# Team and franchiese names

Based on data on standings

```{r}
team_franchise <- readRDS(here::here("data", "standings.rds")) %>%
  distinct(
    season_from,
    season_to,
    team_acronym = purrr::map_chr(strsplit(full_team_name, " "), tail, 1)
  ) %>%
  mutate(
    franchise = case_when(
      # TODO: map season-teams to franchises
      season_to <= 1978 & team_acronym == "Braves" ~ "Clippers",
      season_from >= 2004 & season_to <= 2014 & team_acronym == "Bobcats" ~ "Hornets",

      TRUE ~ team_acronym # else
    )
  )
```

# Data on trades

Process data on trades.

- Assign trades to seasons using April 1 as a cut off date.
- Columns with team names are called `franchise1` and `franchise2` as the names pretend to "season-invariant" names of the franchises, not actual team names used in during seasons.

```{r}
load(here::here("data-src", "data", "NBA_AnalysisData.RData"), 
     envir=NBA_AnalysisData <- new.env())

trades <- NBA_AnalysisData$trade.df_long |>
  as_tibble() |>
  select(-season) |>
  rename(
    franchise1 = P1,
    franchise2 = P2
  ) |>
  mutate(
    season_from = case_when(
      month(Date) >= 4 ~ as.numeric(year(Date)),
      month(Date) <= 3 ~ as.numeric(year(Date)) - 1,
      TRUE ~ as.numeric(NA)
    ),
    season_to = season_from + 1,
    season = paste0(season_from, "-", season_to)
  )

# Make sure of canonical order
stopifnot(with(trades, all(franchise1 < franchise2)))
```

# Divisions

From data on standings extract info on assignment to divisions

- Copy 2018-2019 season to 2019-2020

```{r}
divisions <- readRDS(here::here("data", "standings.rds")) %>%
  distinct(
    season_from,
    division = purrr::map_chr(strsplit(division, " "), 1),
    team_acronym = purrr::map_chr(strsplit(full_team_name, " "), tail, 1)
  )
divisions <- bind_rows(
  divisions,
  divisions %>%
    filter(season_from == 2018) %>%
    mutate(season_from = 2019)
)
```

@@

```{r graphlist}
graphlist <- lapply(
  sort(unique(trades$season_from)),
  function(s) {
    # Select trades
    e <- trades %>%
      filter(season_from == s) %>%
      select(P1, P2)
    # Nodes
    n <- data.frame(
      name = sort(unique(c(e$P1, e$P2)))
    ) %>%
      left_join(
        divisions %>%
          filter(season_from == s) %>%
          rename(div = division),
        by = c("name" = "team_acronym")
      )
    g <- graph_from_data_frame(e, vertices = n, directed=FALSE)
    E(g)$weight <- count_multiple(g)
    g <- simplify(g)
    E(g)$weight <- sqrt(E(g)$weight)
    g$season <- paste0(s, "-", s+1)
    g
  }
)
```
